\section{Environment}

The measurements are captured on a machine with following specifications:

\begin{itemize}
  \item[CPU]{AMD Ryzen 5 1600 @ 3.200\,GHz \\ (6-core, 12-thread processor, 20\,032\,KiB total cache)}
  \item[GPU]{NVIDIA GeForce GTX 1650 Super, 3\,903\,MiB, Driver: 460.56}
  \item[RAM]{32\,GB (4$\times$8\,GB @ 2\,400\,MHz)}
  \item[OS]{Ubuntu 20.04.2 LTS x86\_64}
\end{itemize}

In terms of compiler configuration, \code{g++ 9.3.0} was used as the host compiler and \code{nvcc v11.0.211} was used as the device compiler.

\section{Testing methodology}

GTest, executing tests on different CTU FNSPE systems. Can we prove the implementation is error-proof? Not likely, as the execution is not deterministic. Both the internal state of the node and the results of the operations are tested.

\section{Benchmarking methodology}

After measuring execution times from chosen implementations, we calculate the speedup as the ratio of serial execution time on the CPU and parallel execution time on the GPU.

$$\mathit{Speedup} = \frac{\tau_{\mathit{CPU}}}{\tau_{\mathit{GPU}}}$$

\section{Numerical network microbenchmark}

\todo{Add numerical mesh microbenchmark - numerical networks}

To determine the performance speedup between the GPU and the CPU implementation, a micro-benchmark is introduced with a simple use case. The benchmark consists of incremental inserting and querying unstructured numerical cube mesh.


\section{Comparison against known implementations}

This section describes the comparison of the thesis implementation against other GPU and CPU based implementations. \todo{add cites}

For all of the experiments, 32-bit keys and values are used. Different kinds of pair sequences were used as the input data and each implementation is executed and measured 10 times. Results presented here are averaged. Measurement of CPU based variants are measured using {std::chrono::high\_resolution\_clock::now}, whereas the GPU based ones are measured using CUDA runtime API. The time spent on copying the data from host to device is not included in the measurement.

\todo{Add search benchmark}

\todo{Add insertion benchmark}

\todo{Add an explanation why \cite{awad} is 2-3 times faster than our implementation.} Improved memory access